Mendelian randomization for causal inference in Genome-wide association studies

### 2022-02-16 mr-review.Rmd
added paragraph about zeng_causal_2019-1

@zeng_causal_2019-1

>>>>

Amyotrophic lateral sclerosis is an incurable and largely untreatable neurological condition [@rowland_amyotrophic_2001;@kiernan_amyotrophic_2011]. Working with GWAS summary statistics, @zeng_causal_2019-1 found evidence of a causal relationship between 
type 2 diabetes mellitus and amyotrophic lateral sclerosis in both European and East Asian populations. 
Interestingly, these effects differ in direction; in the European population, type 2 diabetes mellitus 
exerted a neuroprotective effect and diminished the risk of amyotrophic lateral sclerosis, while in the 
East Asian population type 2 diabetes mellitus increased the risk of amyotrophic lateral sclerosis. 
@zeng_causal_2019-1 used inverse variance-weighted methods to estimate the causal effect from summary 
statistics derived from approximately 660,000 type 2 diabetes mellitus European subjects and 81,000 
amyotrophic lateral sclerosis European subjects. Smaller sample sizes (approximately 191,000 type 2 
diabetes mellitus subjects and 4100 amyotrophic lateral sclerosis subjects) were available in the East 
Asian cohorts. The authors used multiple SNP instruments with strong associations with type 2 diabetes 
mellitus [@zeng_causal_2019-1]. The odds ratios of 0.93 and 1.28 in the Europeans and East Asians, 
respectively, were robust against instrument selection, various model misspecifications, and estimation 
biases [@zeng_causal_2019-1]. Egger regression and MR-PRESSO analyses were consistent with the absence of horizonal pleiotropy 
[@zeng_causal_2019-1;@hemani_evaluating_2018;@verbanck_detection_2018;@ong2019implementing]. Future studies may reveal deeper biological insights into these observations and, ultimately, 
contribute to treatments for amyotrophic lateral sclerosis.
